Fair, Stable Exchange Networks through Quenched Merchant Location and Idiosyncratic Trading Costs

### 2019-06-25 writing/Paper/main.tex
Put in edits to paper

Until not too long ago, people used to live in more self-contained villages,
with local markets where people buy different products.  Take clothing as an
example. Most villages had one or several clothing merchants competing for
customers.  There is some healthy competition among merchants within a village,
as people can flexibly switch according to prices and preferences. But
customers would rarely buy clothing from a faraway village, as the transaction
cost associated with visiting that village would be high,  and could not offset
a marginally better price or quality.  On the other hand, people living far
away from any village are disadvantaged as they always incur such high costs.
Similarly, if there is a specialized product that can only be produced in some
villages, it was harder to find customers across village borders,  making it
more difficult for new products and technologies to succeed.

From the industrial to the technological revolution, more efficient
distribution channels have emerged, enabling people to buy almost any product
in online stores.  Independent of a customer's physical location, orders can be
delivered to their door within a day.  More specialized products are now sold
globally, and people living far away from hubs benefit from the availability of
a wide range of products.  While such efficiency is beneficial for the economy
as a whole, it comes at the cost of potential monopolies.  Consider a merchant
that manages to sell a product for a few cents cheaper than its competitors. 
Given modern distribution channels, and ignoring potential differences in
quality, every buyer is now incentivized to buy from that one firm at the
cheaper price. One potential outcome is that the competitors follow along,
increase their efficiency, and a new equilibrium is reached.  However, in a
suboptimal scenario, that one firm leverages the temporarily increased revenue
to further outpace and potentially undercut its competitors.  In the short
term, feedback mechanisms kick in, and an initially minor gap in competitive
advantage, mixed with efficient distribution channels, manifests itself in a suboptimal monopoly.

In this article, we show that such monopolies can be prevented by introducing an idiosyncratic buyer-seller transaction cost.
This tax leverages the benefits of modern distribution channels, making sure that no buyer remains disadvantaged. 
At the same time, innovative products can spread globally, whereas less specialised products remain local. 
This is achieved through the following taxation mechanism: 
Consider a set of $|F|$ firms, all offering the same product, and a set of $|B|$ potential buyers (customers). 
For any buyer $b \in C$ and firm $f \in F$, we define the ordinal distance $D = D(b,f)$ such that the firm $f$ that is physically closest to $b$ has distance $D(b,f)=1$, the next closest has distance $D=2$, and the $n$-th closest has distance $D=n$. 
The tax when customer $b$ buys from the $n$-th closest firm is then proportional to $D^\gamma$, for some exponent $\gamma > 0$. 
If we were to set $\gamma = 0$, everybody can buy from every firm for the same amount of tax. 
If we were to set $\gamma=\infty$, a seller can only buy from its closest firm.
A gamma in-between those two extremes represents a trade-off: 
Close-by products are taxed less than far away products. 
However, because the distance is ordinal, absolute distances do not matter. 
No matter whether the next closest firm is $100$ meters or $100$ kilometres, the taxation is the same. 
In particular, if there is only one firm offering the product, everybody buys for the same tax (since $D=1$ for every buyer), enabling innovations to spread globally.

Through tuning of $\gamma$ and targeted redistribution of tax money to regions of low economic activity, we show that this model is able to fuel a self-consistent economy, 
with fair competition, where innovations can spread freely, but without the chance of monopolies emerging through disproportional feedback mechanisms. 
In summary, we present a model that avoids the winner-take-all monopoly tendency and associated negative externalities of today's internet enabled business through a low cost tax based on a topological constraint.

@@

We simulate an economy of firms and customers in some environment with a distance metric (e.g. a country) and study how firms respond to the idiosyncratic tax.
In this section, we outline the model semi-quantitatively. 
Game-theoretical details are then elaborated in section \ref{sec:game_theory} and concrete applications are described in the subsequent sections. 
Section \ref{sec:conclusions} summarises the results and suggests future directions.

\subsection{Geography}
The geography of our model is one-dimensional with periodic boundaries (i.e. a topological circle) of length $L$. 
Due to the periodic boundary conditions, we measure physical distances with the metric $\min(|b-f|, L-|b-f|)$ for any two points $b$,$f$ (i.e. buyers and firms) on the circle. 
Without loss of generality, we set $L=1$. 
The left-hand-side plots of Figure \ref{fig:fig3x2SingleTimestep} are examples of this `abstract geography'.
Our results not rely on the implemented geography, and any other, potentially more realistic topology could be considered.
However, as we outline above, a simple ring-topology is sufficient for practical purposes, and could be implemented on digital market places such as Amazon.

\subsection{Product}
In a first approach, we consider an economy of only one unique product. 
We assume that all firms produce that same product, but at potentially different production costs. 
Extensions to more products are essentially straight forward and discussed in the outlook section below.

@@

\subsection{Firms}
The firms, like the buyers, are also randomly placed on the circle, and compete non-cooperatively to maximize their profit. 
In this competition, also known as a `game' in the context of gam theory, each firm makes two choices, first capacity (i.e. how much they produce), then price. 
A firm chooses capacity before price because it is difficult to adapt capacity due to supply chain restraints, while it is easier to change a price tag.

@@

\subsection{Idiosyncratic Tax}
The idiosyncratic tax that is introduced by this paper is dependent on a distance measure $D$, or ordinal distance, that gives a discrete distance between every pair of buyers and firms. 
We define the ordinal distance $D$ be defined such that the closest firm to a buyer is at ordinal distance $D=1$, the next closest at $D=2$, and the $n$th closest at $D=n$. 
If buyer $b$ wants to buy a product from its $n$-th closest firms, a tax proportional to $n^\gamma$ for some $\gamma \geqslant 0$ is raised. 
This incentivizes buyers to purchase from more local firms, unless a remote firm offers th same product for a considerably lower price, that offsets the tax.

Crucially, this ordinal distance is calculated for every individual type of product. 
So for instance a car manufacturer and a bakery do not increase each other's distance from their buyers.
Thus, if a firm produces a new product, such as a drug for a previously incurable disease, it is considered at distance $D=1$ from all buyers, at least until a competitor emerges.
This way, new technologies and true innovation can spread freely, whereas more basic products remain more localised.

@@

In this section, we explain how firms set their production quantities and prices based on game theoretically equilibria. 
The reader only interested in qualitative results may skip this section, and continue reading in section \ref{sec:static_game}.

@@

The capacity sub-game acts much like a Cournot game.
\footnote{
A Cournot game is where firms compete only on quantity, and sell all of the quantity at the market clearing price.
} 
The difference is that instead of being assigned a
market clearing price and selling everything they produce, in our model the
firms compete on price in a second sub-game where they choose a price. This
sub-game is necessary because the idiosyncratic tax can provide incentive for
firms to charge different prices. For example, a firm with low costs trying to
sell to non-local buyers would need to undercut firms that ar more local to those potential customers. 
Depending on the tax, this may or may not increase profits.

@@

Note that usually, it is not optimal for a firm to offer products so cheaply that it can sell to the entire market. 
It tends to be more profitable to sell a lower quantity locally at a higher price.

@@

For any buyer $b$ and firm $f$, we define the ordinal distance $D = D(b,f)$
such that the firm $f$ that is physically closest to $b$ has distance
$D(b,f)=1$, the next closest has distance $D=2$, and the $n$th closest has
distance $D=n$.  Thus, if buyer $b$ buys from firm $f$, the total price,
including tax, is given by 
\begin{equation}
	p(b,f) = p_f + p_f ~ r~ \left( D(b,f)^\gamma - 1 \right)
\end{equation}
where $\gamma$ is the coefficient governing how quickly the tax increases, and
$r$ is a simple scaling coefficient (set to $5\%$ in the remainder of this report).

@@

Consider two vending machines right beside one another selling the same soda, with one machine selling soda at a cheaper price than the other. 
Customers will buy from the cheaper machine first until it runs out, and will then switch to the more expensive machine until their demand is satisfied. 
This generic type of consumer behaviour is implemented in our two-step game as well: 
Buyer's start buying from the cheapest buyer first (incl. taxes), then move on to the second cheapest, and so forth. 
Throughout the game, we assume global information, i.e. buyers are perfectly aware of all prices. 
This is good approximation of reality when considering digital market places such as Amazon.

More formally, let the set of buyers be called $B$. 
Each buyer $b\in B$ buys a quantity $q_b =\sum_{f\in F} q_{b,f}$, where $q_{b,f}$ is the amount buyer $b$ buys from the firm $f$. 
The buyer's all have identical demand curves $p_\text{max} + q_b = a$, where $a$ is the endowment, and $p_\text{max}$is the maximum price they are willing to pay (including tax). 
Each buyer tries to maximize $q_b$ while keeping $p_\text{max} + q_b \le a$. 
In other words, they buy from the cheapest available firm until they hit the demand curve and stop buying, or the firm runs out of capacity and the buyer begins buying from the next cheapest, repeating the process.

@@

\subsection{Profit Function}
After having produced $\hat q_f$ units sold at price $p_f$, firm $f$'s profit is given by 
\begin{equation}
    \Pi_f = p_f \sum_{b\in B} q_{b,f} - c_f\hat q_f.
\end{equation}
Also, each $q_{b,f}$ is chosen by the buyer $b$ so that they buy as much as possible while staying under their demand curve $P+Q\leq a$; 
this is where taxes affect the firm's profit.

@@

The firms determine their capacity and price such that they are in a Nash equilibrium. 
Because of the idiosyncratic structure of the game, which depends on the exact location of buyers and sellers, 
it is not possible to determine the Nash equilibrium analytically, for general cases. 
Instead, we use the Gambit library developed by  \cite{gambitPython} to find the Nash equilibrium numerically. 
To this end, the strategy space of capacities and prices are discretised into a sufficiently fine grid. 
Then, the quantity-price two-step game is solved in reverse order: 
For every choice of quantity, the equilibrium price is determined.
Then, the equilibrium quantity is determined. 
Example solutions are provided in the next section.

@@

A different scenario is shown in rows $3$ and $4$ of Figure \ref{fig:static_game}, where the distance-tax applies ($\gamma = 1/2$).
In row 3, we witness again a symmetric outcome, as $c_1 = c_2 = 100$. 
In row 4, we assume again that firm $1$'s production costs are just $1\%$ below firm $2$'s. 
Here, however, the resulting difference in profit is reduced to $11\%$ (and could be reduced further if we were to increase $\gamma$  above $1/2$). 
This shows clearly the balancing effect that the taxation based distance has.

@@

This demonstrates the subtle but important difference induced by our distance-based tax:
A minor difference in production cost, associated with an essentially negligible, random difference in one firm's competitiveness, is dampened to a balanced difference in profits. 
On the other hand, a major difference in production cost, associated with major difference in one firm's competitiveness, is rewarded accordingly by significantly increased profits.

@@

In the previous subsection, we have demonstrated how a distance-tax-free economy disproportionally rewards essentially random, minor differences in firm competitiveness. 
But even more concerningly, the problem does not stay static: 
This disproportional revenue can be used to further invest in aggressive marketing, thereby increasing the difference between the two firms even more. 
This may ultimately run a firm out of business, although it was initially as competitive as the other firms.

>>>>

Until not too long ago, people used to live in more self-contained villages,
with local markets where people buy different products.  Take clothing as an
example. Most villages had one or more clothing merchants competing for
customers.  There is some healthy competition among merchants within a village,
as people can flexibly switch according to prices and preferences. But customers
would rarely buy clothing from a faraway village, as the transaction cost
associated with visiting that village would be high,  and could not offset a
marginally better price or quality.  On the other hand, people living far away
from any village were disadvantaged as they would always incur such high costs.
Similarly, if there is a specialized product that can only be produced in some
villages, it was harder to find customers across village borders, making it more
difficult for new products and technologies to succeed.

From the industrial to the technological revolution, more efficient distribution
channels have emerged, enabling people to buy almost any product in online
stores. Now, orders can be delivered to their door within a day independent of a
customer's physical location  More specialized products are now sold globally,
and even people living far away from cities benefit from the availability of a
wide range of products. While such efficiency is beneficial for the economy as a
whole, it comes at the cost of potential monopolies.  Consider a merchant that
manages to sell a product for a few cents cheaper than its competitors.  Given
modern distribution channels, and ignoring potential differences in quality,
every buyer is now incentivized to buy from that one firm offering the cheaper
price. One potential outcome is that the competitors follow along, increase
their efficiency, and a new equilibrium is reached.  However, in a suboptimal
scenario, that one firm leverages the temporarily increased revenue to further
outpace and potentially undercut its competitors.  In the short term, feedback
mechanisms kick in, and an initially minor gap in competitive advantage, mixed
with efficient distribution channels, manifests itself in a suboptimal monopoly.

In this article, we show that such monopolies can be prevented by introducing an idiosyncratic buyer-seller transaction cost.
This tax leverages the benefits of modern distribution channels, making sure that no buyer remains disadvantaged. 
At the same time, innovative products can spread globally, whereas less specialized products remain local. 
This is achieved through the following taxation mechanism: 
Consider a set of $|F|$ firms, all offering the same product, and a set of $|B|$ potential buyers (customers). 
For any buyer $b \in B$ and firm $f \in F$, we define the ordinal distance $D = D(b,f)$ such that the firm $f$ that is physically closest to $b$ has distance $D(b,f)=1$, the next closest has distance $D=2$, and the $n$-th closest has distance $D=n$. 
The tax when customer $b$ buys from the $n$-th closest firm is then proportional to $D^\gamma$, for some exponent $\gamma > 0$. 
If we were to set $\gamma = 0$, everybody can buy from every firm at single tax rate. 
If we were to set $\gamma=\infty$, a seller can only buy from its closest firm.
A gamma in-between those two extremes represents a trade-off: 
Close-by products are taxed less than far away products. 
However, because the distance is ordinal, absolute distances do not matter. 
No matter whether the next closest firm is $100$ meters or $100$ kilometres, the taxation is the same. 
In particular, if there is only one firm offering the product, everybody buys for the same tax (since $D=1$ for every buyer), enabling innovations to spread globally.

Through tuning of $\gamma$ and targeted redistribution of tax money to regions of low economic activity, we show that this model is able to fuel a self-consistent economy, 
with fair competition, where innovations can spread freely, but without the chance of monopolies emerging through disproportional feedback mechanisms. 
In summary, we present a model that avoids the winner-take-all monopoly tendency and associated negative externalities of today's Internet enabled business through a tax based on a topological constraint \green{the tax is not always small, and definitely not in our examples}.

@@

We simulate an economy of firms and customers in some environment with a distance metric (e.g. a country using road distance) and study how firms respond to the idiosyncratic tax.
In this section, we outline the model semi-quantitatively. 
Game-theoretical details are then elaborated in section \ref{sec:game_theory} and concrete applications are described in the subsequent sections. 
Section \ref{sec:conclusions} summarises the results and suggests future directions.

\subsection{Geography}
The geography of our model is one-dimensional with periodic boundaries (i.e. a topological circle) of length $L$. 
Due to the periodic boundary conditions, we measure physical distances with the metric $\min(|b-f|, L-|b-f|)$ for any two points $b$,$f$ (i.e. buyers and firms) on the circle. 
Without loss of generality, we set $L=1$. 
The left-hand-side plots of Figure \ref{fig:static_game} are examples of this `abstract geography'.
Our results not rely on the implemented geography, and any other, potentially more realistic topology could be considered.
However, as we outline above, a simple ring-topology is sufficient for practical purposes, and could be implemented on digital market places such as Amazon.

\subsection{Product}
In a first approach, we consider an economy of only one product. 
We assume that all firms produce that same product, but at potentially different production costs. 
Extensions to more products are essentially straight forward and discussed in the outlook section below.

@@

\subsection{Firms}
The firms, like the buyers, are also randomly placed on the circle, and compete non-cooperatively to maximize their profit. 
In this competition, also known as a `game' in the context of game theory, each firm makes two choices, first capacity (i.e. how much they produce), then price. 
A firm chooses capacity before price because it is difficult to adapt capacity due to supply chain restraints, while it is easier to change a price tag.

@@

\subsection{Idiosyncratic Tax}
The idiosyncratic tax that is introduced by this paper is dependent on a distance measure $D$, or ordinal distance, that gives a discrete distance between every pair of buyers and firms. 
We define the ordinal distance $D$ be defined such that the closest firm to a buyer is at ordinal distance $D=1$, the next closest at $D=2$, and the $n$th closest at $D=n$. 
If buyer $b$ wants to buy a product from its $n$-th closest firms, a tax proportional to $n^\gamma$ for some $\gamma \geqslant 0$ is raised. 
This incentivizes buyers to purchase from more local firms, unless a remote firm offers the same product for a considerably lower price, offsetting the tax.

Crucially, this ordinal distance is calculated for every individual type of product. 
For instance, a car manufacturer and a bakery do not increase each other's distance from their buyers.
Thus, if a firm produces a new product, such as a drug for a previously incurable disease, it is considered at distance $D=1$ from all buyers, at least until a competitor emerges.
This way, new technologies and true innovation can spread freely, whereas more basic products remain more localized.

@@

In this section, we explain how firms set their production quantities and prices based on game theoretic equilibria. 
The reader only interested in qualitative results may skip this section and continue reading in section \ref{sec:static_game}.

@@

%put the '\footnote' right next to the period to avoid weird spacing between footnote and end of sentence
The capacity sub-game acts much like a Cournot game.\footnote{
A Cournot game is where firms compete only on quantity, and sell all of the quantity at the market clearing price.} 
The difference is that instead of firms being assigned a
market clearing price and selling everything they produce, in our model the
firms compete on price in a second sub-game where they choose a price. This
sub-game is necessary because the idiosyncratic tax can provide incentive for
firms to charge different prices. For example, a firm with low costs trying to
sell to non-local buyers would need to undercut firms that are more local to those potential customers. 
Depending on the tax, this may or may not increase profits.

@@

Note that usually, it is not optimal for a firm to offer products so cheaply that it can sell to the entire market. 
It tends to be more profitable to sell a lower quantity locally at a higher price.\green{I don't know if I would go this far here.}

@@

\green{This is the third time we define ordinal distance. We should cut at least
one}
For any buyer $b$ and firm $f$, we define the ordinal distance $D = D(b,f)$
such that the firm $f$ that is physically closest to $b$ has distance
$D(b,f)=1$, the next closest has distance $D=2$, and the $n$th closest has
distance $D=n$.  Thus, if buyer $b$ buys from firm $f$, the total price,
including tax, is given by 
\begin{equation}
	p(b,f) = p_f + p_f ~ r~ \left( D(b,f)^\gamma - 1 \right)
\end{equation}
where $\gamma$ is the coefficient governing how quickly the tax increases, and
$r$ is a simple scaling coefficient (set to $100\%$ in the remainder of this
report\green{in teh bar plots, $r=1$}).

@@

Consider two vending machines right beside one another selling the same soda, with one machine selling soda at a cheaper price than the other. 
Customers will buy from the cheaper machine first until it runs out, and will then switch to the more expensive machine until their demand is satisfied. 
This generic type of consumer behaviour is implemented in our two-step game as well: 
buyer's start buying from the cheapest buyer first (incl. taxes), then move on to the second cheapest, and so forth. 
Throughout the game, we assume global information, i.e. buyers are perfectly aware of all prices. 
This is good approximation of reality when considering digital market places such as Amazon.

More formally, let the set of buyers be called $B$. 
Each buyer $b\in B$ buys a quantity $q_b =\sum_{f\in F} q_{b,f}$, where $q_{b,f}$ is the amount buyer $b$ buys from the firm $f$. 
The buyers all have identical demand curves $p_\text{max} + q_b = a$, where $a$ is the endowment, and $p_\text{max}$ is the maximum price they buy product at (including tax). 
Each buyer tries to maximize $q_b$ while keeping $p_\text{max} + q_b \le a$. 
In other words, they buy from the cheapest available firm until they hit the demand curve and stop buying, or the firm runs out of capacity and the buyer begins buying from the next cheapest, repeating the process.

@@

\subsection{Profit Function}
After having produced $\hat q_f$ units sold at price $p_f$, firm $f$'s profit is given by 
\begin{equation}
    \Pi_f = p_f \sum_{b\in B} q_{b,f} - c_f\hat q_f,
\end{equation} 
where  $p_f \sum_{b\in B} q_{b,f}$ is the firm's revenue, and $ c_f\hat q_f$ is the firm's total cost.
Taxes affect the firm's profit by altering buyer's preferences, hence changing $q_{b,f}$.

@@

The firms choose their capacity and price such that they are in a Nash equilibrium. 
Because of the idiosyncratic structure of the game, which depends on the exact location of buyers and sellers, 
it is not possible to determine the Nash equilibrium analytically, for general cases. 
Instead, we use the Gambit library developed by  \cite{gambitPython} to find the Nash equilibrium numerically. 
To this end, the strategy space of capacities and prices are discretised into a sufficiently fine grid. 
Then, the quantity-price two-step game is solved in reverse order: 
For every choice of quantity, the equilibrium price is determined.
Then, the equilibrium quantity is determined. 
Example solutions are provided in the next section.

@@

A different scenario is shown in rows $3$ and $4$ of Figure \ref{fig:static_game}, where the distance-tax applies ($\gamma = 1/2$).
In row 3, we again witness a symmetric outcome, as $c_1 = c_2 = 100$. 
In row 4, we assume again that firm $1$'s production costs are just $1\%$ below firm $2$'s. 
Here, however, the resulting difference in profit is reduced to $11\%$ (and could be reduced further if we were to increase $\gamma$  above $1/2$). 
This shows clearly the balancing effect that the taxation based distance has.

@@

This demonstrates the subtle but important difference induced by our distance-based tax:
a minor difference in production cost, associated with an essentially negligible, random difference in one firm's competitiveness, is dampened to a balanced difference in profits. 
On the other hand, a major difference in production cost, associated with major difference in one firm's competitiveness, is rewarded accordingly by significantly increased profits.

@@

In the previous subsection, we have demonstrated how a distance-tax-free economy disproportionally rewards essentially random, minor differences in firm competitiveness. 
But even more concerningly, the problem does not stay static: 
This disproportional revenue can be used to further invest in aggressive marketing, thereby increasing the difference between the two firms even more. 
This may ultimately run a firm out of business, despite initially being as competitive as the other firms.
